---
title: "Homework 2"
author: "Iris (Ziyi) Wang"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

First, define a path to the dataset. 

```{r}
path_to_data = "./data/Trash-Wheel-Collection-Totals-7-2020-1.xlsx"
```


Read the Mr. Trashwheel dataset. 

```{r}
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```

Read precipitation data! For 2018 and 2017. 

```{r}
precip_2018 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-7-2020-1.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)

precip_2017 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-7-2020-1.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Now combine annual precipitation dataframes. In the following code chunk, I create a "helper" tibble that contains pairs of numeric and character ways of representing month, and then merge that (using month number as a key) with the precipitation dataset. This technique is one I use often when I need to recode a moderate or large number of values for a variable. 

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)

precip_df = 
	bind_rows(precip_2018, precip_2017)

precip_df =
	left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

## Problem 2

First, define a path to Problem 2 dataset. 

```{r}
path_to_P2_data = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv"
```


Read the NYC Transit dataset, only retain line, station name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance from the dataset.

```{r}
nyc_subway_df = read_csv(path_to_P2_data) %>% 
    select(c("Line", "Station Name", "Station Latitude", "Station Longitude", 
             "Route1", "Route2", "Route3", "Route4", "Route5", "Route6", 
             "Route7", "Route8", "Route9", "Route10", "Route11", 
             "Entrance Type", "Entry", "Vending", "ADA")) %>% 
	janitor::clean_names() %>% 
    mutate(entry = recode(entry, "YES" = T, "NO" = F),
           vending = recode(vending, "YES" = T, "NO" = F),
           across(route1:route11,as.character))
   
```
The variables in the NYC Transit dataset, called **nyc_subway_df** are: line, station name, station latitude, station longitude, routes served (1-11), entrance type, entry, vending, and ADA (compliance).  
For my data cleaning step, I have changed all the variables name to be more readable and easy to work with (all lowercased, use "_" instead of " "). I also convert the entry and vending variables from character to a logical variable, and convert all routes to be character.
The dimension of resulting dataset **nyc_subway_df** is `r nrow(nyc_subway_df)` x `r ncol(nyc_subway_df)`.  
The dataset is not very tidy to me. It will be better to have a "route_number" column along with a "route_name" column instead of route 1 to route 11.

In this dataset:  
* There are `r nrow(distinct(nyc_subway_df, station_name,line))` distinct stations.  
* If considering distinct stations, then `r nrow(distinct(filter(nyc_subway_df,ada==T), station_name,line))` stations are ADA compliant. Otherwise, `r nrow(filter(nyc_subway_df,ada==T))` stations are ADA compliant.  
* The proportion of station entrances / exits without vending allow entrance is `r nrow(filter(nyc_subway_df,vending == F,entry==T))/nrow(filter(nyc_subway_df,vending == F))`.  

nrow(distinct(filter(nyc_subway_df,vending==T), station_name,line))

Reformat data so that route number and route name are distinct variables, then only keep the A train subdata. 
```{r}
reformated_nyc_subway_df = pivot_longer(nyc_subway_df, 
                                        route1:route11,
                                        names_to = "route_name",
                                        values_to = "route_number") %>% 
    filter(route_number == "A")
```

* There are `r nrow(distinct(reformated_nyc_subway_df, station_name,line))` distinct stations serve the A train.
* `r  nrow(distinct(filter(reformated_nyc_subway_df, ada==T), station_name,line))` of the stations that serve the A train is ADA compliant.


## Problem 3
First, clean the data in pols-month.csv.

```{r}
# month_df = tibble(month = 1:12,month_name = month.name)

pols = read_csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
    separate(col = mon, 
             into = c("year", "month", "day"), 
             sep = "-") %>% 
    mutate(month = as.integer(month),
           month = month.name[month],
           president = prez_gop,
           president = recode(president, "1" = "gop", "0" = "dem")) %>% 
    select(-prez_dem, -prez_gop,-day) %>% 
    arrange(year,month)


```
Second, clean the data in snp.csv.

```{r}
snp = read_csv("./fivethirtyeight_datasets/snp.csv") %>% 
    separate(col = date, 
             into = c("month", "day","year"), 
             sep = "/") %>% 
    mutate(month = as.integer(month),
           month = month.name[month]) %>% 
    select(-day) %>% 
    relocate("year", "month") %>% 
    arrange(year,month)
```

Third, tidy the unemployment data so that it can be merged with the previous datasets.

```{r}
month_abb = 
	tibble(
		month = month.abb,
		month_int = 1:12
	)

unemployment = read_csv("./fivethirtyeight_datasets/unemployment.csv") %>% 
    pivot_longer(Jan:Dec,
                 names_to = "month",
                 values_to = "%unemployment") %>% 
    left_join(month_abb, by = "month") %>% 
    mutate(month_int = month.name[month_int],
           Year = as.character(Year)) %>% 
    select(-month) %>% 
    rename(month = month_int,
           year = Year) %>% 
    relocate(year, month) %>% 
    arrange(year,month)

```

Join the datasets by merging snp into pols, and merging unemployment into the result.
```{r}
result = left_join(left_join(pols,snp),unemployment)
```


The dataset _**pols**_ has dimension of `r nrow(pols)` x `r ncol(pols)`. It shows the number of national politicians who are democratic or republican at any given time. Then range of years is `r min(range(pull(pols,year)))` - `r max(range(pull(pols,year)))`. The key variables in the _**pols**_ are: `r names(pols)`.

The dataset _**snp**_ has dimension of `r nrow(snp)` x `r ncol(snp)`. It shows the closing values of the S&P stock (Standard & Poor’s) index, which is often used as a representative measure of stock market as a whole. Then range of years is `r min(range(pull(snp,year)))` - `r max(range(pull(snp,year)))`. The key variables in the _**snp**_ are: `r names(snp)`.

The dataset _**unemployment**_ has dimension of `r nrow(unemployment)` x `r ncol(unemployment)`. It shows the percentage of unemployment on the associated date. Then range of years is `r min(range(pull(unemployment,year)))` - `r max(range(pull(unemployment,year)))`. The key variables in the _**unemployment**_ are: `r names(unemployment)`.

The resulting dataset, _**result**_, has dimension  is `r nrow(result)` x `r ncol(result)`. Then range of years is `r min(range(pull(result,year)))` - `r max(range(pull(result,year)))`. The key variables in the _**result**_ are: `r names(result)`.

###### Detailed explanation for each variables: 
* year: the year of the measurements  
* month: the month of the measurements  
* gov_gop: the number of republican governors on the associated date  
* sen_gop: the number of republican senators on the associated date  
* rep_gop: the number of republican representatives on the associated date  
* gov_dem: the number of democratic governors on the associated date  
* sen_dem: the number of democratic senators on the associated date  
* rep_dem: the number of democratic representatives on the associated date  
* close: the closing values of the S&P stock index on the associated date  
* %unemployment: percentage of unemployment on the associated date  
